### 2019-09-27 draft/preprint_JCAP_v0.2/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
implemented more changes from revision to v0.1

\documentclass[a4paper]{article}
\usepackage[left=2cm, right=2cm, top=3cm, bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\definecolor{midgreen}{rgb}{0.0,0.675,0.0}
\newcommand{\cgg}[1]{\textcolor{midgreen}{#1}}

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, David Alonso and Emilio Bellini}

\begin{document}
\maketitle
We have carefully addressed the referee's questions and concerns. The
modifications we have performed, for each of the numbered requests, are as
follow:
\begin{enumerate}
  \item  Abstract. Line 9. Added explicit reference to CMB. We have changed
    \textit{expanding on existing algorithms} to \textit{expanding on existing algorithms \cgg{for CMB}}.
  \item  Page 8. After Eq. 2.44. Added GPE2 reference.
  \item  Changed limits of Figure 6 and 7.
  \item  Last sentence of 3.3. Added \textit{or computed exactly as in [41,
      44]}. In the second to last paragraph of Section 4, we added at the end
    \textit{or exactly following Refs.~[41, 44]}.
  \item  In Section 3.3, before previous sentence, we changed \textit{If accurate
    covariances bars} to \textit{If accurate covariance matrix \cgg{elements}}.
\item  In Section 3.4, before Eq. 3.8, I have mentioned the Gaussian
  approximation of the likelihood: \textit{Assuming flat priors \cgg{and a
      fiducial Gaussian likelihood approximation [21]}}.
  \item  Second paragraph of Section 3.4. Changed \textit{for of two
      parameters} to \textit{for the \cgg{two} parameters}.
  \item  At the end of the first paragraph of Section 3.4,  after Eq. 3.8, I
    have added \textit{\cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
        are not being taking into account with this approximation. They,
        however, have a lower statistical weight and will contribute less to
        the posterior distribution.}
}

\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item  
  \item In Section 4, removed the statements about the higher complexity of LSS
    masks.
  \item  In the second paragraph of Section 4, we have added a comment on the
    effect of "real-wold" complications: \textit{With these
    approximation\cgg{s}, computationally speaking, the problem of estimating a
    covariance matrix is as complex as that of computing the power spectrum
    itself, and scales with the number of pixels in the map as $N_{\rm
      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
      extra complications, as position-dependent noise, on their performance.}
}
\end{enumerate}

Extra changes:
\begin{enumerate}
  \item  Changed $\Delta x^2$ to $\Delta^2 x$.
  \item  Before Eq. 2.12, added a comment on how the true power spectra must
    not change much over a bandpower. It says now \textit{Let us denote a given
          bandpower by its index $q$. \cgg{The number of bandpowers and their
            width are determined by the true underlying power spectra, which
            must not vary much within the bandpower}.}
  \item  Introduction. Line 3. Updated the set of references for CMB.
  \item  Added HEALPix project url, reference and credits in the
    Acknowledgments.
\end{enumerate}

\end{document}