### 2019-11-25 draft/preprint_JCAP_v0.2.1-for-arXiv/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
accepted version of paper in JCAP for arXiv without coloured changes

\documentclass[a4paper]{article}
\usepackage[left=2cm, right=2cm, top=3cm, bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\definecolor{midgreen}{rgb}{0.0,0.675,0.0}
\newcommand{\cgg}[1]{\textcolor{midgreen}{#1}}

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, David Alonso and Emilio Bellini}

\begin{document}
\maketitle
We would like to thank the referee for their careful reading of our manuscript.
We have addressed the points raised as described below. All modifications are
highlighted in red in the resubmitted version of the manuscript.

With regards to the specific points raised by the referee:
\begin{enumerate}
  \item We have added an explicit reference to the CMB expertise in the abstract.
  \item We have added a citation to GPE2 after Eq. 2.44.
  \item We have reduced the range of Figures 6 and 7 to highlight the differences in the correlation matrices.
  \item We have included a reference to GPE2 in the context of low-ell covariance estimation.
  \item Thanks for noticing this. This is a typo (left over from a previous version of this sentence that read ``error bars'' instead of ``covariance''). We have changed this sentence as suggested.
  \item We have added the {\sl fiducial Gaussian likelihood} caveat in Section 3.4. We are aware of the biases associated with varying the covariance at each point in parameter space.
  \item Thanks for noticing this typo. This has been fixed.
  \item We have added a caveat about the appropriateness of the Gaussian approximation. We have also noted that the approximate likelihood of Hamimeche \& Lewis also require an accurate evaluation of the power spectrum covariance, and therefore these results are still valuable in that case.
\end{enumerate}

Regarding the additional points raised by the referee:
\begin{enumerate}
 \item We have changed $\Delta x^2$ to $\Delta^2 x$ ($\Delta x^2$ made sense in the case of flat skies, where this is just $\Delta x\,\Delta y$).
 \item We have added a comment about bandpower averaging before Eq. 2.12.
\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item We have expanded the list of references to the CMB legacy for covariance matrix estimation in the introduction, by including a reference to GPE2, to the Planck 2013, 2015 and 2018 likelihood papers, and to Challinog \& Chon 2005 (we also cite the latter in section 2 in the context of mask gradients).
  \item We have toned down the statement about the higher complexity of LSS masks.
  \item We have added a statement about the role of anisotropic noise and other complications at the end of Section 4.
\end{enumerate}

Other modifications:
\begin{enumerate}
  \item We have updated the references to CMB in line 3 of the introduction.
  \item We have added a citation to the HEALPix URL and the corresponding references and acknowledgements.
\end{enumerate}

\end{document}